---
title: "microbenchmark-snippets"
author: "Andrzej Oleś"
output: 
  md_document:
    toc: true
---

```{r microbenchmark, echo=FALSE}
library(microbenchmark)

options(width=100L)
```

## `data.frame` column names {#dataframe-column-names}

Use `names` not `colnames` to access column names in `data.frames`.

```{r data.frame}
microbenchmark(names(mtcars), colnames(mtcars))
identical(names(mtcars), colnames(mtcars))
```

Or even better, use the list representation:

```{r data.frame-list}
list = as.list(mtcars)
microbenchmark(names(list), names(mtcars))
identical(names(list), names(mtcars))
```

## Add elements to list

Use double brackets when adding/setting elements of list.

```{r set-list-el}
l = list()

microbenchmark(
  {l["a"] = 1},
  {l[["a"]] = 1}
)
```

## Coerce result of `strsplit` to a vector

Take the first element of the list rather than unlisting.

```{r strsplit}
x = paste(letters[1:3], collapse = ", ")

microbenchmark(
  unlist(strsplit(x, ", ", fixed=TRUE)),
  strsplit(x, ", ", fixed=TRUE)[[1L]]
)
```

## `sprintf` vs `paste`

`sprintf` can be up to 2x faster.

```{r sprintf}
ext = "pdf"

microbenchmark(
    paste(".*\\.", ext, "$", sep=""),
    paste0(".*\\.", ext, "$"),
    sprintf(".*\\.%s$", ext)
)
```

## `sapply(..., USE.NAMES=FALSE)` vs. `unlist(lapply(...))`

```{r sapply_lapply}
x = letters

microbenchmark(
  sapply(x, nchar, USE.NAMES=FALSE),
  unlist(lapply(x, nchar))
)
```

## Interleaving two vectors

The fastest approach is to use "the `rbind` trick".

```{r interleave}
length = 1000L
a <- rep_len(letters, length)
b <- rep_len(LETTERS, length)

r <- function(a, b) {
  as.vector(rbind(a, b))
}

f <- function(a, b) {
  v = vector(mode = "character", length = 2L*length)
  idx <- 2L*1:length
  v[idx-1L] = a
  v[idx] = b
  v
}

g <- function(a, b) {
  idx <- order(c(seq_along(a), seq_along(b)))
  c(a,b)[idx]
}

h <- function(a, b) {
  unlist(mapply(function(x, y) c(x, y), a, b, SIMPLIFY=FALSE, USE.NAMES=FALSE))
}

microbenchmark(
  r(a, b),
  f(a, b),
  g(a, b),
  h(a, b)
)
```

## Insert elements into a vector

Fill-in a preallocated results vector.

```{r insert}
n = 1000
lines = sample(letters, n, TRUE)
idx = sort(sample(n, size = 10))

f = function(lines, idx, txt = "whatever") {
  idx_length = length(idx)
  lines_length = length(lines)
  v = vector(mode = "character", length = lines_length+idx_length)
  idx <- idx + seq_len(idx_length)
  v[-idx] <- lines
  v[idx] <- txt
  v
}

g = function(lines, idx, txt = "whatever") {
  v = rep_len(NA_character_, length(lines))
  v[idx] = txt
  v = as.vector(rbind(lines, v))
  v[!is.na(v)]
}

identical(f(lines, idx), g(lines, idx))

microbenchmark(
  f(lines, idx),
  g(lines, idx)
)
```

## `tabulate` as a much faster alternative to `table`

For programming it typically makes more sense to call `tabulate` on a well formed integer vector. If possible, specify `nbins`.

```{r tabulate}
library(EBImage)
x = readImage(system.file('images', 'shapes.png', package='EBImage'))
x = x[110:512,1:130]
y = bwlabel(x)
nbins = max(y)

microbenchmark(
  table(y),
  tabulate(y),
  tabulate(y, nbins)
)
```

## Session info

```{r sessionInfo}
sessionInfo()
```
