---
title: "microbenchmark-snippets"
author: "Andrzej Oleś"
date: "`r doc_date()`"
output: BiocStyle::md_document
---

```{r microbenchmark, echo=FALSE}
library(microbenchmark)

options(width=100L)
```

## `data.frame` column names {#dataframe-column-names}

Use `names` not `colnames` to access column names in `data.frames`.

```{r data.frame}
microbenchmark(names(mtcars), colnames(mtcars))
identical(names(mtcars), colnames(mtcars))
```

Or even better, use the list representation:

```{r data.frame-list}
list = as.list(mtcars)
microbenchmark(names(list), names(mtcars))
identical(names(list), names(mtcars))
```

## Add elements to list

Use double brackets when adding/setting elements of list.

```{r set-list-el}
list = list()
microbenchmark({list["a"] = 1}, {list[["a"]] = 1})
```

## Coerce result of `strsplit` to a vector

Take the first element of the list rather than unlisting.

```{r strsplit}
x = paste(letters[1:3], collapse = ", ")

microbenchmark(
  unlist(strsplit(x, ", ", fixed=TRUE)),
  strsplit(x, ", ", fixed=TRUE)[[1L]]
)
```

## `sprintf` vs `paste`

`sprintf` can be almost 2x faster~

```{r sprintf}
ext = "pdf"

microbenchmark(
    paste(".*\\.", ext, "$", sep=""),
    paste0(".*\\.", ext, "$"),
    sprintf(".*\\.%s$", ext)
)
```

## `sapply(..., USE.NAMES=FALSE)` vs. `unlist(lapply(...))`

```{r sapply_lapply}
x = letters
microbenchmark(
  sapply(x, nchar, USE.NAMES=FALSE),
  unlist(lapply(x, nchar))
)
```

## Session info

```{r sessionInfo}
sessionInfo()
```
